---
title: "Annotations plots"
format: html
editor: visual
---

## **Packages**

```{r}
library(RColorBrewer)
library(viridis)
library(tidyverse)
library(broom)
library(mgcv)
library(ggeffects)
library(patchwork)
library(paletteer)
library(ggthemes)
```

## **Loading Data**

### **Annotation datasets**

```{r}

setwd("C:/Users/TACNZ4/OneDrive - cchmc/Brazil seasonality data")

meantemp= read.table("monthly temperature.txt", header=TRUE)

meanprecip= read.table("monthly precipitation.txt", header=TRUE)

daylength=read.table("daylength minutes.txt", header=TRUE)

school= read.table("school breaks.txt", header=TRUE)
```

### **Datasus datasets**

```{r}
setwd("C:/Users/TACNZ4/OneDrive - cchmc/Datasus seasonality")
dados.sp=read_rds("dadossp.rds")
dados.pe=read_rds("dadospe.rds")
dados.df=read_rds("dadosdf.rds")
dados.rs=read_rds("dadosrs.rds")
```

## **Preparing data**

### **Ordering months (July -\>June) - Environmental Variables**

```{r}


meantemp$Month=factor(meantemp$Month, levels=c(7,8,9,10,11,12,1,2,3,4,5,6))
meanprecip$Month=factor(meanprecip$Month, levels=c(7,8,9,10,11,12,1,2,3,4,5,6))
daylength$Month=factor(daylength$Month, levels=c(7,8,9,10,11,12,1,2,3,4,5,6))
school$Month=factor(school$Month, levels=c(7,8,9,10,11,12,1,2,3,4,5,6))


```

### **Annotations by location**

```{r}
#São paulo
tempsp <- meantemp|>
  select(Month, SP)
precipsp <-  meanprecip|>
  select(Month, SP)
daysp <-  daylength|>
  select(Month, SP)
schoolsp <-  school|>
  select(Month, SP)

#Porto Alegre
temppoa <- meantemp|>
  select(Month, POA)
precippoa <-  meanprecip|>
  select(Month, POA)
daypoa <-  daylength|>
  select(Month, POA)
schoolpoa <-  school|>
  select(Month, POA)

#Brasilia
tempbra <- meantemp|>
  select(Month, Brasilia)
precipbra <-  meanprecip|>
  select(Month, Brasilia)
daybra <-  daylength|>
  select(Month, Brasilia)
schoolbra <-  school|>
  select(Month, Brasilia)

#Recife
temprec <- meantemp|>
  select(Month, Recife)
preciprec <-  meanprecip|>
  select(Month, Recife)
dayrec <-  daylength|>
  select(Month, Recife)
schoolrec <-  school|>
  select(Month, Recife)

```

### **Separating dataSUS by location, age and emergencies**

```{r}

#### São paulo
### Only São Paulo municipality (code 355030)
spdata= dados.sp %>%
  subset(MUNIC_MOV==c("355030")) 
### filter 0-18 y.o.
## COD_IDADE is the measuring unit for age. 4=years, 3=months, 2=days
# IDADE is age
spdata= spdata %>%
  subset((COD_IDADE=="4" & IDADE <=18)|(COD_IDADE=="3"))
### Only emergencies
## CAR_INT is the type of admission. "02" is urgency/emergency
spdata= spdata %>%
  subset(CAR_INT==c("02"))
## Selecting relevant variables:
spdata<- spdata|>
  select(MES_CMPT, ANO_CMPT, DIAG_PRINC,DT_INTER, DT_SAIDA, MUNIC_MOV)


#### Porto Alegre
### Only Porto Alegre municipality (code 431490)
poadata= dados.rs %>%
  subset(MUNIC_MOV==c("431490")) 
### filter 0-18 y.o.
## COD_IDADE is the measuring unit for age. 4=years, 3=months, 2=days
# IDADE is age
poadata= poadata %>%
  subset((COD_IDADE=="4" & IDADE <=18)|(COD_IDADE=="3"))
### Only emergencies
## CAR_INT is the type of admission. "02" is urgency/emergency
poadata= poadata %>%
  subset(CAR_INT=="02")
## Selecting relevant variables:
poadata<- poadata|>
  select(MES_CMPT, ANO_CMPT, DIAG_PRINC,DT_INTER, DT_SAIDA, MUNIC_MOV)


#### Brasilia
### Only Brasilia municipality (code 530010)
bradata= dados.df %>%
  subset(MUNIC_MOV==c("530010")) 
### filter 0-18 y.o.
## COD_IDADE is the measuring unit for age. 4=years, 3=months, 2=days
# IDADE is age
bradata= bradata%>%
  subset((COD_IDADE=="4" & IDADE <=18)|(COD_IDADE=="3"))
### Only emergencies
## CAR_INT is the type of admission. "02" is urgency/emergency
bradata= bradata %>%
  subset(CAR_INT=="02")
## Selecting relevant variables:
bradata<- bradata|>
  select(MES_CMPT, ANO_CMPT, DIAG_PRINC,DT_INTER, DT_SAIDA, MUNIC_MOV)

#### Recife
### Only Recife municipality (code 530010)
recdata= dados.pe %>%
  subset(MUNIC_MOV==c("261160")) 
### filter 0-18 y.o.
## COD_IDADE is the measuring unit for age. 4=years, 3=months, 2=days
# IDADE is age
recdata= recdata%>%
  subset((COD_IDADE=="4" & IDADE <=18)|(COD_IDADE=="3"))
### Only emergencies
## CAR_INT is the type of admission. "02" is urgency/emergency
recdata= recdata %>%
  subset(CAR_INT=="02")
## Selecting relevant variables:
recdata<- recdata|>
  select(MES_CMPT, ANO_CMPT, DIAG_PRINC,DT_INTER, DT_SAIDA, MUNIC_MOV)

```

### **Subsetting Asthma**

```{r}
### Asthma ICDs - Choose all ICDs that have "J45"
## DIAG_PRINC is the ICD10 code for the main diagnosis at admission

# São Paulo
asthmasp= spdata %>%
  subset(str_detect(DIAG_PRINC, "J45"))

# Porto Alegre
asthmapoa= poadata %>%
  subset(str_detect(DIAG_PRINC, "J45"))

# Brasilia
asthmabra= bradata %>%
  subset(str_detect(DIAG_PRINC, "J45"))

# Recife
asthmarec= recdata %>%
  subset(str_detect(DIAG_PRINC, "J45"))

```

### **Subsetting Mental Health**

```{r}
### Mental Health ICDs - Choose all ICDs that have "F"

# São Paulo
mentalsp= spdata %>%
  subset(str_detect(DIAG_PRINC, "F"))

# Porto Alegre
mentalpoa= poadata %>%
  subset(str_detect(DIAG_PRINC, "F"))

# Brasilia
mentalbra= bradata %>%
  subset(str_detect(DIAG_PRINC, "F"))

# Recife
mentalrec= recdata %>%
  subset(str_detect(DIAG_PRINC, "F"))

```

## **Creating the environmental annotation plots**

### **São Paulo**

```{r}
# Temperature plot
ptempSP= ggplot(tempsp, aes(x = factor(Month), y = 1, fill=SP)) + 
  geom_tile() +
  scale_fill_paletteer_c("ggthemes::Orange-Blue Diverging", direction=-1, limits=c(14,28))+
  #scale_fill_viridis(option="turbo")+
  scale_x_discrete(breaks = seq(1,12, by = 1),
                   labels = c("jan", "feb", "mar", "apr", "may", "jun", "jul",
                              "aug", "sep", "oct", "nov", "dec"),
                   guide = guide_axis(n.dodge = 2)) +
  guides(fill=guide_legend(title="Temperature (oC)"))+
labs(x = "month", y = "Temperature") +
  theme_void()+
  theme(legend.position = "right",
        legend.title = element_text(),
        legend.text = element_text(), 
        axis.title.x = element_text(),
        axis.text.x = element_text(),
        axis.title.y = element_text())
ptempSP 

# Precipitation plot
pprecipSP= ggplot(precipsp, aes(x = factor(Month), y = 1, fill=SP)) + 
  geom_tile() +
  scale_fill_gradient(low= "white", high="darkseagreen4", limits=c(0,400))+
  labs(x = " ", y = "Precipitation") +
  guides(fill=guide_legend(title="Precipitation (mm)"))+
  theme_void()+ 
  theme(legend.title = element_text(), axis.title.y = element_text())


pprecipSP 


# Daylength plot
pdaySP= ggplot(daysp, aes(x = factor(Month), y = 1, fill=SP)) + 
  geom_tile() +
  scale_fill_paletteer_c("ggthemes::Orange Light", direction=1, limits=c(610,850))+
  labs(x = " ", y = "Daylength") +
  guides(fill=guide_legend(title="Daylenght (min)"))+
  theme_void()+ 
  theme(legend.title = element_text(), axis.title.y = element_text())
pdaySP 


# School breaks plot
pschoolSP= ggplot(schoolsp, aes(x = factor(Month), y = 1, fill=SP)) + 
  geom_tile() +
  scale_fill_grey(start = 1, end = 0.4)+
  scale_x_discrete(breaks = seq(1,12, by = 1),
                   labels = c("jan", "feb", "mar", "apr", "may", "jun", "jul",
                              "aug", "sep", "oct", "nov", "dec"),
                   guide = guide_axis(n.dodge = 2)) +
  guides(fill=guide_legend(title="School Breaks"))+
  labs(x = " ", y = "School Breaks") +
  theme_void()+ 
  theme(legend.title = element_text(), axis.title.y = element_text())
pschoolSP
```

### **Porto Alegre**

```{r}
# Temperature plot
ptemppoa= ggplot(temppoa, aes(x = factor(Month), y = 1, fill=POA)) + 
  geom_tile() +
  scale_fill_paletteer_c("ggthemes::Orange-Blue Diverging", direction=-1, limits=c(14,28))+
  #scale_fill_viridis(option="turbo")+
  scale_x_discrete(breaks = seq(1,12, by = 1),
                   labels = c("jan", "feb", "mar", "apr", "may", "jun", "jul",
                              "aug", "sep", "oct", "nov", "dec"),
                   guide = guide_axis(n.dodge = 2)) +
  guides(fill=guide_legend(title="Temperature (oC)"))+
labs(x = "month", y = "Temperature") +
  theme_void()+
  theme(legend.position = "right",
        legend.title = element_text(),
        legend.text = element_text(), 
        axis.title.x = element_text(),
        axis.text.x = element_text(),
        axis.title.y = element_text())
ptemppoa 

# Precipitation plot
pprecippoa= ggplot(precippoa, aes(x = factor(Month), y = 1, fill=POA)) + 
  geom_tile() +
  scale_fill_gradient(low= "white", high="darkseagreen4", limits=c(0,400))+
  #scale_fill_viridis(option="brewer blues")+
  scale_x_discrete(breaks = seq(1,12, by = 1),
                   labels = c("jan", "feb", "mar", "apr", "may", "jun", "jul",
                              "aug", "sep", "oct", "nov", "dec"),
                   guide = guide_axis(n.dodge = 2)) +
  labs(x = " ", y = "Precipitation") +
  guides(fill=guide_legend(title="Precipitation (mm)"))+
  theme_void()+ 
  theme(legend.title = element_text(), axis.title.y = element_text())


pprecippoa 


# Daylength plot
pdaypoa= ggplot(daypoa, aes(x = factor(Month), y = 1, fill=POA)) + 
  geom_tile() +
  scale_fill_paletteer_c("ggthemes::Orange Light", direction=1, limits=c(610,850))+
  labs(x = " ", y = "Daylength") +
  guides(fill=guide_legend(title="Daylenght (min)"))+
  theme_void()+ 
  theme(legend.title = element_text(), axis.title.y = element_text())
pdaypoa 


# School breaks plot
pschoolpoa= ggplot(schoolpoa, aes(x = factor(Month), y = 1, fill=POA)) + 
  geom_tile() +
  scale_fill_grey(start = 1, end = 0.4)+
  scale_x_discrete(breaks = seq(1,12, by = 1),
                   labels = c("jan", "feb", "mar", "apr", "may", "jun", "jul",
                              "aug", "sep", "oct", "nov", "dec"),
                   guide = guide_axis(n.dodge = 2)) +
  guides(fill=guide_legend(title="School Breaks"))+
  labs(x = " ", y = "School Breaks") +
  theme_void()+ 
  theme(legend.title = element_text(), axis.title.y = element_text())
pschoolpoa
```

### **Brasilia**

```{r}
# Temperature plot
ptempbra= ggplot(tempbra, aes(x = factor(Month), y = 1, fill=Brasilia)) + 
  geom_tile() +
  scale_fill_paletteer_c("ggthemes::Orange-Blue Diverging", direction=-1, limits=c(14,28))+
  #scale_fill_viridis(option="turbo")+
  scale_x_discrete(breaks = seq(1,12, by = 1),
                   labels = c("jan", "feb", "mar", "apr", "may", "jun", "jul",
                              "aug", "sep", "oct", "nov", "dec"),
                   guide = guide_axis(n.dodge = 2)) +
  guides(fill=guide_legend(title="Temperature (oC)"))+
labs(x = "month", y = "Temperature") +
  theme_void()+
  theme(legend.position = "right",
        legend.title = element_text(),
        legend.text = element_text(), 
        axis.title.x = element_text(),
        axis.text.x = element_text(),
        axis.title.y = element_text())
ptempbra 

# Precipitation plot
pprecipbra= ggplot(precipbra, aes(x = factor(Month), y = 1, fill=Brasilia)) + 
  geom_tile() +
  scale_fill_gradient(low= "white", high="darkseagreen4", limits=c(0,400))+
  #scale_fill_viridis(option="brewer blues")+
  scale_x_discrete(breaks = seq(1,12, by = 1),
                   labels = c("jan", "feb", "mar", "apr", "may", "jun", "jul",
                              "aug", "sep", "oct", "nov", "dec"),
                   guide = guide_axis(n.dodge = 2)) +
  labs(x = " ", y = "Precipitation") +
  guides(fill=guide_legend(title="Precipitation (mm)"))+
  theme_void()+ 
  theme(legend.title = element_text(), axis.title.y = element_text())


pprecipbra


# Daylength plot
pdaybra= ggplot(daybra, aes(x = factor(Month), y = 1, fill=Brasilia)) + 
  geom_tile() +
  scale_fill_paletteer_c("ggthemes::Orange Light", direction=1, limits=c(610,850))+
  labs(x = " ", y = "Daylength") +
  guides(fill=guide_legend(title="Daylenght (min)"))+
  theme_void()+ 
  theme(legend.title = element_text(), axis.title.y = element_text())
pdaybra 


# School breaks plot
pschoolbra= ggplot(schoolbra, aes(x = factor(Month), y = 1, fill=Brasilia)) + 
  geom_tile() +
  scale_fill_grey(start = 1, end = 0.4)+
  scale_x_discrete(breaks = seq(1,12, by = 1),
                   labels = c("jan", "feb", "mar", "apr", "may", "jun", "jul",
                              "aug", "sep", "oct", "nov", "dec"),
                   guide = guide_axis(n.dodge = 2)) +
  guides(fill=guide_legend(title="School Breaks"))+
  labs(x = " ", y = "School Breaks") +
  theme_void()+ 
  theme(legend.title = element_text(), axis.title.y = element_text())
pschoolbra


```

### **Recife**

```{r}
# Temperature plot
ptemprec= ggplot(temprec, aes(x = factor(Month), y = 1, fill=Recife)) + 
  geom_tile() +
  scale_fill_paletteer_c("ggthemes::Orange-Blue Diverging", direction=-1, limits=c(14,28))+
  #scale_fill_viridis(option="turbo")+
  scale_x_discrete(breaks = seq(1,12, by = 1),
                   labels = c("jan", "feb", "mar", "apr", "may", "jun", "jul",
                              "aug", "sep", "oct", "nov", "dec"),
                   guide = guide_axis(n.dodge = 2)) +
  guides(fill=guide_legend(title="Temperature (oC)"))+
labs(x = "month", y = "Temperature") +
  theme_void()+
  theme(legend.position = "right",
        legend.title = element_text(),
        legend.text = element_text(), 
        axis.title.x = element_text(),
        axis.text.x = element_text(),
        axis.title.y = element_text())
ptemprec

# Precipitation plot
ppreciprec= ggplot(preciprec, aes(x = factor(Month), y = 1, fill=Recife)) + 
  geom_tile() +
  scale_fill_gradient(low= "white", high="darkseagreen4", limits=c(0,400))+
  #scale_fill_viridis(option="brewer blues")+
  scale_x_discrete(breaks = seq(1,12, by = 1),
                   labels = c("jan", "feb", "mar", "apr", "may", "jun", "jul",
                              "aug", "sep", "oct", "nov", "dec"),
                   guide = guide_axis(n.dodge = 2)) +
  labs(x = " ", y = "Precipitation") +
  guides(fill=guide_legend(title="Precipitation (mm)"))+
  theme_void()+ 
  theme(legend.title = element_text(), axis.title.y = element_text())


ppreciprec


# Daylength plot
pdayrec= ggplot(dayrec, aes(x = factor(Month), y = 1, fill=Recife)) + 
  geom_tile() +
  scale_fill_paletteer_c("ggthemes::Orange Light", direction=1, limits=c(610,850))+
 # scale_x_discrete(breaks = seq(1,12, by = 1),
  #                 labels = c("jan", "feb", "mar", "apr", "may", "jun", "jul",
   #                           "aug", "sep", "oct", "nov", "dec"),
    #               guide = guide_axis(n.dodge = 2)) +
  labs(x = " ", y = "Daylength") +
  guides(fill=guide_legend(title="Daylenght (min)"))+
  theme_void()+ 
  theme(legend.title = element_text(), axis.title.y = element_text())
pdayrec 


# School breaks plot
pschoolrec= ggplot(schoolrec, aes(x = factor(Month), y = 1, fill=Recife)) + 
  geom_tile() +
  scale_fill_grey(start = 1, end = 0.4)+
  scale_x_discrete(breaks = seq(1,12, by = 1),
                   labels = c("jan", "feb", "mar", "apr", "may", "jun", "jul",
                              "aug", "sep", "oct", "nov", "dec"),
                   guide = guide_axis(n.dodge = 2)) +
  guides(fill=guide_legend(title="School Breaks"))+
  labs(x = " ", y = "School Breaks") +
  theme_void()+ 
  theme(legend.title = element_text(), axis.title.y = element_text())
pschoolrec
```

## **Creating the average admission plots - Asthma**

### **São Paulo**

```{r}
freqssp <- asthmasp |> 
  mutate(across(.cols = c('MES_CMPT'), as.factor)) |> 
  group_by(MUNIC_MOV, MES_CMPT, .drop = FALSE)|> 
  summarise(n = n())|>
  ungroup() |> 
  mutate(MES_CMPT= as.numeric(MES_CMPT)) |> 
  left_join(days <- tibble(
    MES_CMPT = seq(1, 12, 1),
    DAYS = c(31, 28.25, 31, 30, 31, 30, 31, 31, 30, 31, 30, 31),
    ADJUST = max(DAYS)/DAYS), by = 'MES_CMPT') |> 
  mutate(DAY_AVG_ADM = round(n/DAYS/12, digits = 3), # Avg per day
         MONTH_AVG_ADM = round(n*ADJUST/12, digits = 2), # Avg per month
         MONTH_COUNT_ADM = round(n*ADJUST, digits = 0)) # Total per month

freqssp$MES_CMPT=factor(freqssp$MES_CMPT, levels=c(7,8,9,10,11,12,1,2,3,4,5,6))

pSP <- ggplot(freqssp) + 
  geom_bar(aes(MES_CMPT, weight= MONTH_AVG_ADM, alpha=.6), color="gray") +
  scale_size_manual(values = c(0.75, 0.15), guide = 'none') +
  ylim(0,350)+
  labs(x = "month", y = "average monthly admissions") +
  ggtitle("São Paulo")+
  theme_bw() +
  theme(legend.position = 'none',
        panel.grid.major = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 15),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.margin = margin(b = -7, l = -40),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        axis.ticks.x = element_blank())+
 annotate("text", x = Inf, y = Inf, label = c("PTR=1.72"), hjust = 1.1, vjust = 2, size = 4)
pSP
```

### **Porto Alegre**

```{r}
freqspoa <- asthmapoa |> 
  mutate(across(.cols = c('MES_CMPT'), as.factor)) |> 
  group_by(MUNIC_MOV, MES_CMPT, .drop = FALSE)|> 
  summarise(n = n())|>
  ungroup() |> 
  mutate(MES_CMPT= as.numeric(MES_CMPT)) |> 
  left_join(days <- tibble(
    MES_CMPT = seq(1, 12, 1),
    DAYS = c(31, 28.25, 31, 30, 31, 30, 31, 31, 30, 31, 30, 31),
    ADJUST = max(DAYS)/DAYS), by = 'MES_CMPT') |> 
  mutate(DAY_AVG_ADM = round(n/DAYS/12, digits = 3), # Avg per day
         MONTH_AVG_ADM = round(n*ADJUST/12, digits = 2), # Avg per month
         MONTH_COUNT_ADM = round(n*ADJUST, digits = 0)) # Total per month

freqspoa$MES_CMPT=factor(freqssp$MES_CMPT, levels=c(7,8,9,10,11,12,1,2,3,4,5,6))

pPOA <- ggplot(freqspoa) + 
  geom_bar(aes(MES_CMPT, weight= MONTH_AVG_ADM, alpha=.6), color="gray") +
  scale_size_manual(values = c(0.75, 0.15), guide = 'none') +
  labs(x = "month", y = "average monthly admissions") +
    ylim(0,230)+
  ggtitle("Porto Alegre")+
  theme_bw() +
  theme(legend.position = 'none',
        panel.grid.major = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 15),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.margin = margin(b = -7, l = -40),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        axis.ticks.x = element_blank())+
   annotate("text", x = Inf, y = Inf, label = c("PTR=3.71"), hjust = 1.1, vjust = 2, size = 4)

pPOA
```

### **Brasilia**

```{r}
freqsbra <- asthmabra |> 
  mutate(across(.cols = c('MES_CMPT'), as.factor)) |> 
  group_by(MUNIC_MOV, MES_CMPT, .drop = FALSE)|> 
  summarise(n = n())|>
  ungroup() |> 
  mutate(MES_CMPT= as.numeric(MES_CMPT)) |> 
  left_join(days <- tibble(
    MES_CMPT = seq(1, 12, 1),
    DAYS = c(31, 28.25, 31, 30, 31, 30, 31, 31, 30, 31, 30, 31),
    ADJUST = max(DAYS)/DAYS), by = 'MES_CMPT') |> 
  mutate(DAY_AVG_ADM = round(n/DAYS/12, digits = 3), # Avg per day
         MONTH_AVG_ADM = round(n*ADJUST/12, digits = 2), # Avg per month
         MONTH_COUNT_ADM = round(n*ADJUST, digits = 0)) # Total per month

freqsbra$MES_CMPT=factor(freqsbra$MES_CMPT, levels=c(7,8,9,10,11,12,1,2,3,4,5,6))

pBRA <- ggplot(freqsbra) + 
  geom_bar(aes(MES_CMPT, weight= MONTH_AVG_ADM, alpha=.6), color="gray") +
  scale_size_manual(values = c(0.75, 0.15), guide = 'none') +
  labs(x = "month", y = "average monthly admissions") +
  ggtitle("Brasilia")+
  theme_bw() +
  theme(legend.position = 'none',
        panel.grid.major = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 15),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.margin = margin(b = -7, l = -40),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        axis.ticks.x = element_blank())+
    ylim(0,120)+
     annotate("text", x = Inf, y = Inf, label = c("PTR=2.09"), hjust = 1.1, vjust = 2, size = 4)

pBRA
```

### **Recife**

```{r}
freqsrec <- asthmarec |> 
  mutate(across(.cols = c('MES_CMPT'), as.factor)) |> 
  group_by(MUNIC_MOV, MES_CMPT, .drop = FALSE)|> 
  summarise(n = n())|>
  ungroup() |> 
  mutate(MES_CMPT= as.numeric(MES_CMPT)) |> 
  left_join(days <- tibble(
    MES_CMPT = seq(1, 12, 1),
    DAYS = c(31, 28.25, 31, 30, 31, 30, 31, 31, 30, 31, 30, 31),
    ADJUST = max(DAYS)/DAYS), by = 'MES_CMPT') |> 
  mutate(DAY_AVG_ADM = round(n/DAYS/12, digits = 3), # Avg per day
         MONTH_AVG_ADM = round(n*ADJUST/12, digits = 2), # Avg per month
         MONTH_COUNT_ADM = round(n*ADJUST, digits = 0)) # Total per month

freqsrec$MES_CMPT=factor(freqsrec$MES_CMPT, levels=c(7,8,9,10,11,12,1,2,3,4,5,6))

pREC <- ggplot(freqsrec) + 
  geom_bar(aes(MES_CMPT, weight= MONTH_AVG_ADM, alpha=.6), color="gray") +
  scale_size_manual(values = c(0.75, 0.15), guide = 'none') +
  labs(x = "month", y = "average monthly admissions") +
  ggtitle("Recife")+
  theme_bw() +
  theme(legend.position = 'none',
        panel.grid.major = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 15),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.margin = margin(b = -7, l = -40),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        axis.ticks.x = element_blank())+
    ylim(0,140)+
     annotate("text", x = Inf, y = Inf, label = c("PTR=1.38"), hjust = 1.1, vjust = 2, size = 4)
pREC

```

## **Creating the average admission plots - Mental health**

### **São Paulo**

```{r}
freqssp.mental <- mentalsp |> 
  mutate(across(.cols = c('MES_CMPT'), as.factor)) |> 
  group_by(MUNIC_MOV, MES_CMPT, .drop = FALSE)|> 
  summarise(n = n())|>
  ungroup() |> 
  mutate(MES_CMPT= as.numeric(MES_CMPT)) |> 
  left_join(days <- tibble(
    MES_CMPT = seq(1, 12, 1),
    DAYS = c(31, 28.25, 31, 30, 31, 30, 31, 31, 30, 31, 30, 31),
    ADJUST = max(DAYS)/DAYS), by = 'MES_CMPT') |> 
  mutate(DAY_AVG_ADM = round(n/DAYS/12, digits = 3), # Avg per day
         MONTH_AVG_ADM = round(n*ADJUST/12, digits = 2), # Avg per month
         MONTH_COUNT_ADM = round(n*ADJUST, digits = 0)) # Total per month

freqssp.mental$MES_CMPT=factor(freqssp.mental$MES_CMPT, levels=c(7,8,9,10,11,12,1,2,3,4,5,6))

pSP.mental <- ggplot(freqssp.mental) + 
  geom_bar(aes(MES_CMPT, weight= MONTH_AVG_ADM, alpha=.6), color="gray") +
  scale_size_manual(values = c(0.75, 0.15), guide = 'none') +
  labs(x = "month", y = "average monthly admissions") +
  ggtitle("São Paulo")+
  theme_bw() +
  theme(legend.position = 'none',
        panel.grid.major = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 15),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.margin = margin(b = -7, l = -40),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        axis.ticks.x = element_blank())+
    ylim(0,65)+
     annotate("text", x = Inf, y = Inf, label = c("PTR=1.21"), hjust = 1.1, vjust = 2, size = 4)

pSP.mental
```

### **Porto Alegre**

```{r}
freqspoa.mental <- mentalpoa |> 
  mutate(across(.cols = c('MES_CMPT'), as.factor)) |> 
  group_by(MUNIC_MOV, MES_CMPT, .drop = FALSE)|> 
  summarise(n = n())|>
  ungroup() |> 
  mutate(MES_CMPT= as.numeric(MES_CMPT)) |> 
  left_join(days <- tibble(
    MES_CMPT = seq(1, 12, 1),
    DAYS = c(31, 28.25, 31, 30, 31, 30, 31, 31, 30, 31, 30, 31),
    ADJUST = max(DAYS)/DAYS), by = 'MES_CMPT') |> 
  mutate(DAY_AVG_ADM = round(n/DAYS/12, digits = 3), # Avg per day
         MONTH_AVG_ADM = round(n*ADJUST/12, digits = 2), # Avg per month
         MONTH_COUNT_ADM = round(n*ADJUST, digits = 0)) # Total per month

freqspoa.mental$MES_CMPT=factor(freqspoa.mental$MES_CMPT, levels=c(7,8,9,10,11,12,1,2,3,4,5,6))

pPOA.mental <- ggplot(freqspoa.mental) + 
  geom_bar(aes(MES_CMPT, weight= MONTH_AVG_ADM, alpha=.6), color="gray") +
  scale_size_manual(values = c(0.75, 0.15), guide = 'none') +
  labs(x = "month", y = "average monthly admissions") +
  ggtitle("Porto Alegre")+
  theme_bw() +
  theme(legend.position = 'none',
        panel.grid.major = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 15),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.margin = margin(b = -7, l = -40),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        axis.ticks.x = element_blank())+
    ylim(0,45)+
     annotate("text", x = Inf, y = Inf, label = c("PTR=1.22"), hjust = 1.1, vjust = 2, size = 4)

pPOA.mental
```

### **Brasilia**

```{r}
freqsbra.mental <- mentalbra |> 
  mutate(across(.cols = c('MES_CMPT'), as.factor)) |> 
  group_by(MUNIC_MOV, MES_CMPT, .drop = FALSE)|> 
  summarise(n = n())|>
  ungroup() |> 
  mutate(MES_CMPT= as.numeric(MES_CMPT)) |> 
  left_join(days <- tibble(
    MES_CMPT = seq(1, 12, 1),
    DAYS = c(31, 28.25, 31, 30, 31, 30, 31, 31, 30, 31, 30, 31),
    ADJUST = max(DAYS)/DAYS), by = 'MES_CMPT') |> 
  mutate(DAY_AVG_ADM = round(n/DAYS/12, digits = 3), # Avg per day
         MONTH_AVG_ADM = round(n*ADJUST/12, digits = 2), # Avg per month
         MONTH_COUNT_ADM = round(n*ADJUST, digits = 0)) # Total per month

freqsbra.mental$MES_CMPT=factor(freqsbra.mental$MES_CMPT, levels=c(7,8,9,10,11,12,1,2,3,4,5,6))

pBRA.mental <- ggplot(freqsbra.mental) + 
  geom_bar(aes(MES_CMPT, weight= MONTH_AVG_ADM, alpha=.6), color="gray") +
  scale_size_manual(values = c(0.75, 0.15), guide = 'none') +
  labs(x = "month", y = "average monthly admissions") +
  ggtitle("Brasilia")+
  theme_bw() +
  theme(legend.position = 'none',
        panel.grid.major = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 15),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.margin = margin(b = -7, l = -40),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        axis.ticks.x = element_blank())+
    ylim(0,30)+
     annotate("text", x = Inf, y = Inf, label = c("PTR=1.23"), hjust = 1.1, vjust = 2, size = 4)

pBRA.mental
```

### **Recife**

```{r}
freqsrec.mental <- mentalrec |> 
  mutate(across(.cols = c('MES_CMPT'), as.factor)) |> 
  group_by(MUNIC_MOV, MES_CMPT, .drop = FALSE)|> 
  summarise(n = n())|>
  ungroup() |> 
  mutate(MES_CMPT= as.numeric(MES_CMPT)) |> 
  left_join(days <- tibble(
    MES_CMPT = seq(1, 12, 1),
    DAYS = c(31, 28.25, 31, 30, 31, 30, 31, 31, 30, 31, 30, 31),
    ADJUST = max(DAYS)/DAYS), by = 'MES_CMPT') |> 
  mutate(DAY_AVG_ADM = round(n/DAYS/12, digits = 3), # Avg per day
         MONTH_AVG_ADM = round(n*ADJUST/12, digits = 2), # Avg per month
         MONTH_COUNT_ADM = round(n*ADJUST, digits = 0)) # Total per month

freqsrec.mental$MES_CMPT=factor(freqsrec.mental$MES_CMPT, levels=c(7,8,9,10,11,12,1,2,3,4,5,6))

pREC.mental <- ggplot(freqsrec.mental) + 
  geom_bar(aes(MES_CMPT, weight= MONTH_AVG_ADM, alpha=.6), color="gray") +
  scale_size_manual(values = c(0.75, 0.15), guide = 'none') +
  labs(x = "month", y = "average monthly admissions") +
  ggtitle("Recife")+
  theme_bw() +
  theme(legend.position = 'none',
        panel.grid.major = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 15),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.margin = margin(b = -7, l = -40),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        axis.ticks.x = element_blank())+
    ylim(0,23)+
     annotate("text", x = Inf, y = Inf, label = c("PTR=1.47"), hjust = 1.1, vjust = 2, size = 4)

pREC.mental
```

## **Creating patches for annotations**

```{r}
# São Paulo
patchsp= (pschoolSP +  
            theme (axis.title.y = element_blank(), legend.position = "none"))/
  (pdaySP +  
     theme (axis.title.y = element_blank(), legend.position = "none"))/
  (pprecipSP +  
     theme (axis.title.y = element_blank(), legend.position = "none"))/
  ( ptempSP +  
      theme (axis.title.y = element_blank(), legend.position = "none")) 

# Porto Alegre
patchpoa= (pschoolpoa +  
            theme (legend.position = "none"))/
  (pdaypoa +  
     theme (legend.position = "none"))/
  (pprecippoa +  
     theme (legend.position = "none"))/
  ( ptemppoa +  
      theme (legend.position = "none")) 

# Brasilia
patchbra= (pschoolbra +  
            theme (axis.title.y = element_blank(), legend.position = "none"))/
  (pdaybra +  
     theme (axis.title.y = element_blank(), legend.position = "none"))/
  (pprecipbra +  
     theme (axis.title.y = element_blank(), legend.position = "none"))/
  ( ptempbra +  
      theme (axis.title.y = element_blank(), legend.position = "none")) 

# Recife
patchrec= (pschoolrec +  
            theme (axis.title.y = element_blank()))/
  (pdayrec +  
     theme (axis.title.y = element_blank()))/
  (ppreciprec +  
     theme (axis.title.y = element_blank()))/
  ( ptemprec +  
      theme (axis.title.y = element_blank())) 


```

## **Assembling Asthma plots**

```{r}
((pPOA/
    patchpoa +  plot_layout(heights = c(2, 1))) |
((pSP+  theme (axis.title.y = element_blank()))/
       patchsp  + plot_layout(heights = c(2, 1))) |
((pBRA+  theme (axis.title.y = element_blank()))/
       patchbra  + plot_layout(heights = c(2, 1))) |
((pREC+  theme (axis.title.y = element_blank()))/
       patchrec  + plot_layout(heights = c(2, 1), guides="collect"))) + plot_annotation(title = 'Asthma', theme = theme(plot.title = element_text(size = 16) ))

ggsave("asthma_annotations.jpg", width = 12, height = 7)
```

## **Assembling Mental Health plots**

```{r}
((pPOA.mental/
    patchpoa +  plot_layout(heights = c(2, 1))) |
((pSP.mental+  theme (axis.title.y = element_blank()))/
       patchsp  + plot_layout(heights = c(2, 1))) |
((pBRA.mental+  theme (axis.title.y = element_blank()))/
       patchbra  + plot_layout(heights = c(2, 1))) |
((pREC.mental+  theme (axis.title.y = element_blank()))/
       patchrec  + plot_layout(heights = c(2, 1), guides="collect")) ) +
  plot_annotation(title = 'Mental Health', theme = theme(plot.title = element_text(size = 16) ))

ggsave("mental_annotations.jpg", width = 12, height = 7)
```

## **Calculating Effect size**

### **Peak-Minus-Trough**

```{r}

#### Asthma

magsp.asthma=mean(sort(freqssp$MONTH_COUNT_ADM)[11]+sort(freqssp$MONTH_COUNT_ADM)[12])-mean(sort(freqssp$MONTH_COUNT_ADM)[1]+sort(freqssp$MONTH_COUNT_ADM)[2])
#2689

magpoa.asthma=mean(sort(freqspoa$MONTH_COUNT_ADM)[11]+sort(freqspoa$MONTH_COUNT_ADM)[12])-mean(sort(freqspoa$MONTH_COUNT_ADM)[1]+sort(freqspoa$MONTH_COUNT_ADM)[2])
#3247

magbra.asthma=mean(sort(freqsbra$MONTH_COUNT_ADM)[11]+sort(freqsbra$MONTH_COUNT_ADM)[12])-mean(sort(freqsbra$MONTH_COUNT_ADM)[1]+sort(freqsbra$MONTH_COUNT_ADM)[2])
#1213

magrec.asthma=mean(sort(freqsrec$MONTH_COUNT_ADM)[11]+sort(freqsrec$MONTH_COUNT_ADM)[12])-mean(sort(freqsrec$MONTH_COUNT_ADM)[1]+sort(freqsrec$MONTH_COUNT_ADM)[2])
#880


#Adjusted by total admissions
magsp.asthma/sum(freqssp$MONTH_COUNT_ADM)
#0.08352488

magpoa.asthma/sum(freqspoa$MONTH_COUNT_ADM)
#0.1977707

magbra.asthma/sum(freqsbra$MONTH_COUNT_ADM)
#0.1220076

magrec.asthma/sum(freqsrec$MONTH_COUNT_ADM)
#0.05055728



#### Mental Health

magsp.mental=mean(sort(freqssp.mental$MONTH_COUNT_ADM)[11]+sort(freqssp.mental$MONTH_COUNT_ADM)[12])-mean(sort(freqssp.mental$MONTH_COUNT_ADM)[1]+sort(freqssp.mental$MONTH_COUNT_ADM)[2])
#247

magpoa.mental=mean(sort(freqspoa.mental$MONTH_COUNT_ADM)[11]+sort(freqspoa.mental$MONTH_COUNT_ADM)[12])-mean(sort(freqspoa.mental$MONTH_COUNT_ADM)[1]+sort(freqspoa.mental$MONTH_COUNT_ADM)[2])
#182

magbra.mental=mean(sort(freqsbra.mental$MONTH_COUNT_ADM)[11]+sort(freqsbra.mental$MONTH_COUNT_ADM)[12])-mean(sort(freqsbra.mental$MONTH_COUNT_ADM)[1]+sort(freqsbra.mental$MONTH_COUNT_ADM)[2])
#242

magrec.mental=mean(sort(freqsrec.mental$MONTH_COUNT_ADM)[11]+sort(freqsrec.mental$MONTH_COUNT_ADM)[12])-mean(sort(freqsrec.mental$MONTH_COUNT_ADM)[1]+sort(freqsrec.mental$MONTH_COUNT_ADM)[2])
#135


#Adjusted by total admissions
magsp.mental/sum(freqssp.mental$MONTH_COUNT_ADM)
#0.02177952

magpoa.mental/sum(freqspoa.mental$MONTH_COUNT_ADM)
#0.03165606

magbra.mental/sum(freqsbra.mental$MONTH_COUNT_ADM)
#0.03134309

magrec.mental/sum(freqsrec.mental$MONTH_COUNT_ADM)
#0.01705699
```

### **Peak-Trough-Ratio**

ptrpoa.mental

```{r}

### Asthma
ptrsp.asthma=mean(sort(freqssp$MONTH_COUNT_ADM)[11]+sort(freqssp$MONTH_COUNT_ADM)[12])/mean(sort(freqssp$MONTH_COUNT_ADM)[1]+sort(freqssp$MONTH_COUNT_ADM)[2])
#1.718024

ptrpoa.asthma=mean(sort(freqspoa$MONTH_COUNT_ADM)[11]+sort(freqspoa$MONTH_COUNT_ADM)[12])/mean(sort(freqspoa$MONTH_COUNT_ADM)[1]+sort(freqspoa$MONTH_COUNT_ADM)[2])
#3.70809

ptrbra.asthma=mean(sort(freqsbra$MONTH_COUNT_ADM)[11]+sort(freqsbra$MONTH_COUNT_ADM)[12])/mean(sort(freqsbra$MONTH_COUNT_ADM)[1]+sort(freqsbra$MONTH_COUNT_ADM)[2])
#2.086918

ptrrec.asthma=mean(sort(freqsrec$MONTH_COUNT_ADM)[11]+sort(freqsrec$MONTH_COUNT_ADM)[12])/mean(sort(freqsrec$MONTH_COUNT_ADM)[1]+sort(freqsrec$MONTH_COUNT_ADM)[2])
#1.380623

### Mental  Health
ptrsp.mental=mean(sort(freqssp.mental$MONTH_COUNT_ADM)[11]+sort(freqssp.mental$MONTH_COUNT_ADM)[12])/mean(sort(freqssp.mental$MONTH_COUNT_ADM)[1]+sort(freqssp.mental$MONTH_COUNT_ADM)[2])
#1.212199

ptrpoa.mental=mean(sort(freqspoa.mental$MONTH_COUNT_ADM)[11]+sort(freqspoa.mental$MONTH_COUNT_ADM)[12])/mean(sort(freqspoa.mental$MONTH_COUNT_ADM)[1]+sort(freqspoa.mental$MONTH_COUNT_ADM)[2])
#1.224969

ptrbra.mental=mean(sort(freqsbra.mental$MONTH_COUNT_ADM)[11]+sort(freqsbra.mental$MONTH_COUNT_ADM)[12])/mean(sort(freqsbra.mental$MONTH_COUNT_ADM)[1]+sort(freqsbra.mental$MONTH_COUNT_ADM)[2])
# 1.233051

ptrrec.mental=mean(sort(freqsrec.mental$MONTH_COUNT_ADM)[11]+sort(freqsrec.mental$MONTH_COUNT_ADM)[12])/mean(sort(freqsrec.mental$MONTH_COUNT_ADM)[1]+sort(freqsrec.mental$MONTH_COUNT_ADM)[2])
#1.477032
```

**Population size:**

**São Paulo: 11,451,999**

**Porto Alegre: 1.332.845**

**Brasilia: 2.817.381**

**Recife: 1.488.920**
